Limb-darkening and exoplanets II: Choosing the Best Law for Minimum Bias Retrieval of Transit Parameters

### 2016-01-08 first_response_to_referee.txt

Bella Lock (Mrs) 
Assistant Editor, Royal Astronomical Society

Dear Mrs. Bella,

 I am resubmitting the manuscript MNRAS/MN-15-3639-MJ:

 "Limb-darkening and exoplanets II: Choosing the Best 
  Law for Minimum Bias Retrieval of Transit Parameters"

 Authors: Nestor Espinoza & Andres Jordan

 We thank the referee for his/her careful reading of the manuscript. 
 You will find below our detailed response to the referee's comments 
 as indented text after each of his/her comments.

 Besides the changes made in response to the referee's comments we have:

     - Corrected minor typos on the manuscript.

########################################################################################
>> My main suggestion is that the paper would benefit from a more quantitative 
>> description of the trade-off between precision and accuracy. This work indicates 
>> that fitting for two limb darkening coefficients with a judiciously chosen model 
>> can significantly reduce bias. In practice, however, many light curves are not 
>> sufficiently precise to tightly constrain more than one limb darkening parameter. 
>> Adding an additional degree of freedom with a second limb darkening coefficient 
>> may cause the uncertainty on other important parameters, particularly the transit 
>> depth, to blow up. The resulting loss in accuracy may not be worth the improvement 
>> in precision due to choosing a two-parameter model over fitting for  a linear limb 
>> darkening coefficient.
>>
>> One way to address this question is to give guidelines about how precise a 
>> light curve must be to merit adding a second limb darkening parameter to the 
>> model. This could be discussed for a few select examples (e.g., a hot Jupiter 
>> transiting a Sun-like star).  I leave it up to the authors whether/how to address 
>> this point.

RESPONSE

>> It would also be helpful to investigate how much bias is introduced by using a linear 
>> limb darkening law, since this model is commonly used in practice. Again, the authors 
>> can address this at their discretion.

RESPONSE

>> I would recommend adding a few sentences to the abstract to give a little more context 
>> for the paper. For example, something to the effect of: "Very precise measurements of 
>> exoplanet transit light curves enable limb darkening parameters to be fitted empirically, 
>> rather than fixed on theoretical model predictions that may have unknown systematic errors. 
>> However, even fitting for the limb darkening coefficients can introduce bias in the measured 
>> transit parameters depending on the chosen limb darkening model parameterization." I recommend 
>> adding this because because fitting for the coefficients directly is still a relatively new 
>> approach in this field, and it may not be obvious to everyone why you're not just using the 
>> standard nonlinear law.

RESPONSE

>> There are several cases of very precise transit light curves where the limb darkening coefficients 
>> predicted from theoretical models does not provide a good fit to the data (Knutson et al. 2007, ApJ, 
>> 655, 564; Kreidberg et al. 2015, ApJ, 814, 66).  Please cite these in the introduction.

RESPONSE

>> State in the text what time sampling you use for your model light curves, and how sensitive your 
>> results are to the choice of time sampling.

RESPONSE

>> There are a few uneccessary lines in the code run_ld_exosim.py that should be removed:
>>      import pymc
>>      sys.path.append('../utilities/')

RESPONSE

### 2016-01-08 first_response_to_referee.txt

>> It would also be helpful to investigate how much bias is introduced by using a linear 
>> limb darkening law, since this model is commonly used in practice. Again, the authors 
>> can address this at their discretion.

RESPONSE

>> I would recommend adding a few sentences to the abstract to give a little more context 
>> for the paper. For example, something to the effect of: "Very precise measurements of 
>> exoplanet transit light curves enable limb darkening parameters to be fitted empirically, 
>> rather than fixed on theoretical model predictions that may have unknown systematic errors. 
>> However, even fitting for the limb darkening coefficients can introduce bias in the measured 
>> transit parameters depending on the chosen limb darkening model parameterization." I recommend 
>> adding this because because fitting for the coefficients directly is still a relatively new 
>> approach in this field, and it may not be obvious to everyone why you're not just using the 
>> standard nonlinear law.

We thank the referee for this suggestion. We have rewritten the first few sentences in the abstract 
in order to provide the suggested context to the paper.

@@

>> There are several cases of very precise transit light curves where the limb darkening coefficients 
>> predicted from theoretical models does not provide a good fit to the data (Knutson et al. 2007, ApJ, 
>> 655, 564; Kreidberg et al. 2015, ApJ, 814, 66).  Please cite these in the introduction.

We have added the suggested cites to the discussion in the introduction.

@@

>> State in the text what time sampling you use for your model light curves, and how sensitive your 
>> results are to the choice of time sampling.

RESPONSE

>> There are a few uneccessary lines in the code run_ld_exosim.py that should be removed:
>>      import pymc
>>      sys.path.append('../utilities/')

We thank the referee for catching this typo in the code. The lines have been removed from the code.

### 2016-01-08 which_law_2_use.tex
Added responses to some of the referees comments.

\begin{abstract}
The parametrization of the limb-darkening effect of a star has a direct impact 
on the parameters retrieved from transit lightcurves such as the exoplanet's radius, scaled semi-major axis and inclination. Studies of the effect of limb darkening assumptions on the retrieved parameters have mostly focused on the widely used quadratic limb-darkening law, leaving out 
other proposed laws that are in fact better descriptions of model intensity profiles. In this work, we show that 
laws such as the logarithmic, square-root and three-parameter laws do a better job at retrieving the mentioned parameters from transit 
lightcurves than the widely used quadratic law and therefore we recommend using those instead. We also provide code to guide the decision of when to use each of the laws,  which we note has a dependence on both stellar and transit parameters. Finally, we demonstrate that the so-called exponential law is non-physical as it typically produces negative intensities close to the limb and should therefore not be used.
\end{abstract}

@@

In a recent study \citep[][hereon EJ15]{ej2015}, we showed that the accuracy is actually catching up with the precision 
of these measurements due to our imperfect understanding of the limb-darkening effect. In particular, we showed that there are important 
biases both when fixing and fitting for the limb-darkening coefficients in the light curve fitting process. The biases when fitting for the limb darkening coefficients arise from the fact 
that the popular and widely used quadratic law is unable to model the complex intensity profile of real stars,  and the biases when fixing the coefficients arise for the
same reason plus the facts that different methods of fitting the model intensity profiles (including transit light curve versus intensity profile fitting) give rise to different limb-darkening coefficients and 
the fact that we have an imperfect knowledge of the real intensity profiles of stars. We showed that fixing the limb-darkening coefficients is 
the worst option if one aims to obtain precise \textit{and} accurate transit parameters, because the biases arise from three different sources, 
with the last one (the fact that we do not have a perfect modelling of the stellar intensity profiles of real stars) having an unknown but possibly 
large impact.

@@

For the simulations, the applied strategies to sample physically plausible coefficients from two-parameter laws 
are as follows, all of which use numbers $q_1 \in (0,1)$ and $q_2 \in (0,1)$:
\begin{itemize}
\item For the quadratic limb-darkening law, we fit for the numbers $q_1 = (u_1 + u_2)^2$ and $q_2 = u_1/2(u_1+u_2)$. 
To obtain the coefficients from those fitted numbers, we use the relations $u_1 = 2\sqrt{q_1}q_2$ and $u_2 =\sqrt{q_1}(1-2q_2)$, 
which are derived in \cite{kipping2013}.
\item For the square-root law, we fit for the numbers $q_1 = (s_1 + s_2)^2$ and $q_2 = s_2/2(s_1+s_2)$. 
To obtain the coefficients from those fitted numbers, we use the relations $s_1 = \sqrt{q_1}(1-2q_2)$ and $s_2 =2\sqrt{q_1}q_2$, 
which are derived in \cite{kipping2013}.
\item For the logarithmic law, we fit for the numbers $q_1 = (1-l_2)^2$ and $q_2 = (1-l_1)/(1-l_2)$. 
To obtain the coefficients from those fitted numbers, we use the relations $l_1 = 1-\sqrt{q_1}q_2$ and $l_2=1-\sqrt{q_1}$, 
which are derived in \S 2.1.
\end{itemize}
For the three-parameter law, we use the formalism and code described in \cite{kipping2015}.  The code used to generate our simulations is available
in GitHub\footnote{\url{http://www.github.com/nespinoza/ld-simulations/}}, and can be used to perform analyses in a finer grid than the ones 
published here.

@@

The question of when to use each of those laws is an important one. In particular, we have explored the relation between the real, underlying 
transit parameters, the stellar host properties and the bias on the retrieved transit parameters. We have stated no specific rule regarding when to use each 
law, but rather provided only guidelines based on our simulations. This is a choice made on purpose, as the selection of which law to use has to be done in a case-by-case basis, reason for which we provide a code with which one can obtain the results of the same simulations for specific 
cases and with different resolutions\footnote{\url{http://www.github.com/nespinoza/ld-simulations/}}. It is important to note that in this work we have used the 
{\em Kepler} bandpass to study the relation between the biases on the retrieved transit parameters and the stellar host properties, but different bandpasses should give 
rise to  different functional relations because they give rise to different model intensity profiles, which are used in those simulations to approximate the 
expected intensity profile and the adequacy of each limb-darkening law at following it. Limb-darkening coefficients have to be calculated for each 
particular case, which is straightforward to do using the algorithms described in EJ15\footnote{\url{http://www.github.com/nespinoza/limb-darkening/}}.

@@

\begin{thebibliography}{99}
\bibitem[\protect\citeauthoryear{Claret}{2000}]{claret2000} Claret, A., 2000, A\&A, 363, 1081.
\bibitem[\protect\citeauthoryear{Claret \& Hauschildt}{2003}]{claret2003} Claret, A. \& Hauschildt, P. H., 2003, 
A\&A, 412, 241.
\bibitem[\protect\citeauthoryear{D\'iaz-Cordov\'ez \& Gim\'enez}{1992}]{diazgimenez1992} D\'iaz-Cordov\'ez, J.,
\& Gim\'enez, A., 1992, A\&A, 259, 227.
\bibitem[\protect\citeauthoryear{Dorn et al.}{2015}]{dorn2015} Dorn, C., Khan, A., Heng, K., Alibert, Y., Connolly, J.,
Benz, W. \& Tackley, P., 2015, A\&A, 577, 83.
\bibitem[\protect\citeauthoryear{Kipping}{2013}]{kipping2013} Kipping, D. M., 2013, MNRAS, 435, 2152.
\bibitem[\protect\citeauthoryear{Kipping}{2014}]{kipping2014AP} Kipping, D. M., 2014, MNRAS, 440, 2164.
\bibitem[\protect\citeauthoryear{Kipping}{2015}]{kipping2015} Kipping, D. M., 2015, arXiv:1509.03483.
\bibitem[\protect\citeauthoryear{Klinglesmith \& Sobieski}{1970}]{klinglesmith1970} Klinglesmith, D. A. \&
Sobieski, S., 1970, AJ, 75, 175.
\bibitem[\protect\citeauthoryear{Espinoza \& Jord\'an}{2015}]{ej2015} Espinoza, N. \& Jord\'an, A., 2015, 
MNRAS, 450, 1879.
\bibitem[\protect\citeauthoryear{Howarth}{2011}]{howarth2011} Howarth, I., 2011, 
MNRAS, 413, 1515.
\bibitem[\protect\citeauthoryear{Klinglesmith \& Sobieski}{1970}]{klinglesmith1970} Klinglesmith, D. A. \&
Sobieski, S., 1970, AJ, 75, 175.
\bibitem[\protect\citeauthoryear{Koch et~al.}{2010}]{koch2010} Koch et~al. 2010, ApJ, 713, L79
\bibitem[\protect\citeauthoryear{Kreidberg}{2015}]{batman2015} Kreidberg, L., 2015, arXiv:1507.08285.
\bibitem[\protect\citeauthoryear{Mandel \& Agol}{2002}]{ma2002} Mandel, K \& Agol, E., 2002, ApJ,
580, L171.
\bibitem[\protect\citeauthoryear{Ricker et al.}{2014}]{ricker2014} Ricker et al., 2014, SPIE, 9143.
\bibitem[\protect\citeauthoryear{Rowe et al.}{2015}]{rowe2015} Rowe et al., 2015, ApJS, 217, 16.

>>>>

\begin{abstract}
With the advent of very precise measurements of exoplanet transit lightcurves both from ground and space based observatories, it has been 
shown that it is now possible to fit the limb-darkening coefficients in the transit-fitting procedure rather than fix them to theoretical values. In 
addition, this strategy has been shown to give better results as fixing the coefficients to theoretical values can give rise to important yet unknown systematic 
errors, which directly impact the physical properties of the system derived from such lightcurves such as the exoplanet's radius, scaled 
semi-major axis and inclination. However, studies of the effect of limb darkening assumptions on the retrieved parameters 
have mostly focused on the widely used quadratic limb-darkening law, leaving out other proposed laws that are in fact better descriptions 
of model intensity profiles, which in principle could diminish the bias even further. In this work, we show that laws such as the logarithmic, 
square-root and three-parameter do a better job in general at retrieving the mentioned parameters from transit lightcurves than the widely 
used quadratic law and therefore we recommend using those instead. We also provide code to guide the decision of when to use each of 
the laws,  which we note has a dependence on both stellar and transit parameters. Finally, we demonstrate that the so-called exponential 
law is non-physical as it typically produces negative intensities close to the limb and should therefore not be used.
\end{abstract}

@@

In a recent study \citep[][hereon EJ15]{ej2015}, we showed that the accuracy is actually catching up with the precision 
of these measurements due to our imperfect understanding of the limb-darkening effect. In particular, we showed that there are important 
biases both when fixing and fitting for the limb-darkening coefficients in the light curve fitting process. The biases when fitting for the 
limb darkening coefficients arise from the fact that the popular and widely used quadratic law is unable to model the complex intensity profile 
of real stars \citep[which has also been seen on other studies, see, e.g.,][]{knutson2007, kreidberg2015},  and the biases when fixing the coefficients arise for the
same reason plus the facts that different methods of fitting the model intensity profiles (including transit light curve versus intensity profile fitting) give rise to different limb-darkening coefficients and 
the fact that we have an imperfect knowledge of the real intensity profiles of stars. We showed that fixing the limb-darkening coefficients is 
the worst option if one aims to obtain precise \textit{and} accurate transit parameters, because the biases arise from three different sources, 
with the last one (the fact that we do not have a perfect modelling of the stellar intensity profiles of real stars) having an unknown but possibly 
large impact.

@@

For the simulations, the applied strategies to sample physically plausible coefficients from two-parameter laws 
are as follows, all of which use numbers $q_1 \in (0,1)$ and $q_2 \in (0,1)$:
\begin{itemize}
\item For the quadratic limb-darkening law, we fit for the numbers $q_1 = (u_1 + u_2)^2$ and $q_2 = u_1/2(u_1+u_2)$. 
To obtain the coefficients from those fitted numbers, we use the relations $u_1 = 2\sqrt{q_1}q_2$ and $u_2 =\sqrt{q_1}(1-2q_2)$, 
which are derived in \cite{kipping2013}.
\item For the square-root law, we fit for the numbers $q_1 = (s_1 + s_2)^2$ and $q_2 = s_2/2(s_1+s_2)$. 
To obtain the coefficients from those fitted numbers, we use the relations $s_1 = \sqrt{q_1}(1-2q_2)$ and $s_2 =2\sqrt{q_1}q_2$, 
which are derived in \cite{kipping2013}.
\item For the logarithmic law, we fit for the numbers $q_1 = (1-l_2)^2$ and $q_2 = (1-l_1)/(1-l_2)$. 
To obtain the coefficients from those fitted numbers, we use the relations $l_1 = 1-\sqrt{q_1}q_2$ and $l_2=1-\sqrt{q_1}$, 
which are derived in \S 2.1.
\end{itemize}
For the three-parameter law, we use the formalism and code described in \cite{kipping2015}.  The code used to generate our simulations is available
in GitHub\footnote{\url{http://www.github.com/nespinoza/ld-exosim/}}, and can be used to perform analyses in a finer grid than the ones 
published here.

@@

The question of when to use each of those laws is an important one. In particular, we have explored the relation between the real, underlying 
transit parameters, the stellar host properties and the bias on the retrieved transit parameters. We have stated no specific rule regarding when to use each 
law, but rather provided only guidelines based on our simulations. This is a choice made on purpose, as the selection of which law to use has to be done in a case-by-case basis, reason for which we provide a code with which one can obtain the results of the same simulations for specific 
cases and with different resolutions\footnote{\url{http://www.github.com/nespinoza/ld-exosim/}}. It is important to note that in this work we have used the 
{\em Kepler} bandpass to study the relation between the biases on the retrieved transit parameters and the stellar host properties, but different bandpasses should give 
rise to  different functional relations because they give rise to different model intensity profiles, which are used in those simulations to approximate the 
expected intensity profile and the adequacy of each limb-darkening law at following it. Limb-darkening coefficients have to be calculated for each 
particular case, which is straightforward to do using the algorithms described in EJ15\footnote{\url{http://www.github.com/nespinoza/limb-darkening/}}.

@@

\begin{thebibliography}{99}
\bibitem[\protect\citeauthoryear{Claret}{2000}]{claret2000} Claret, A., 2000, A\&A, 363, 1081.
\bibitem[\protect\citeauthoryear{Claret \& Hauschildt}{2003}]{claret2003} Claret, A. \& Hauschildt, P. H., 2003, 
A\&A, 412, 241.
\bibitem[\protect\citeauthoryear{D\'iaz-Cordov\'ez \& Gim\'enez}{1992}]{diazgimenez1992} D\'iaz-Cordov\'ez, J.,
\& Gim\'enez, A., 1992, A\&A, 259, 227.
\bibitem[\protect\citeauthoryear{Dorn et al.}{2015}]{dorn2015} Dorn, C., Khan, A., Heng, K., Alibert, Y., Connolly, J.,
Benz, W. \& Tackley, P., 2015, A\&A, 577, 83.
\bibitem[\protect\citeauthoryear{Kipping}{2013}]{kipping2013} Kipping, D. M., 2013, MNRAS, 435, 2152.
\bibitem[\protect\citeauthoryear{Kipping}{2014}]{kipping2014AP} Kipping, D. M., 2014, MNRAS, 440, 2164.
\bibitem[\protect\citeauthoryear{Kipping}{2015}]{kipping2015} Kipping, D. M., 2015, arXiv:1509.03483.
\bibitem[\protect\citeauthoryear{Klinglesmith \& Sobieski}{1970}]{klinglesmith1970} Klinglesmith, D. A. \&
Sobieski, S., 1970, AJ, 75, 175.
\bibitem[\protect\citeauthoryear{Espinoza \& Jord\'an}{2015}]{ej2015} Espinoza, N. \& Jord\'an, A., 2015, 
MNRAS, 450, 1879.
\bibitem[\protect\citeauthoryear{Howarth}{2011}]{howarth2011} Howarth, I., 2011, 
MNRAS, 413, 1515.
\bibitem[\protect\citeauthoryear{Klinglesmith \& Sobieski}{1970}]{klinglesmith1970} Klinglesmith, D. A. \&
Sobieski, S., 1970, AJ, 75, 175.
\bibitem[\protect\citeauthoryear{Knutson et al.}{2007}]{knutson2007} Knutson, H., Charbonneau, D., Noyes, R., et al., 2007, ApJ, 655, 564.

\bibitem[\protect\citeauthoryear{Koch et~al.}{2010}]{koch2010} Koch et~al. 2010, ApJ, 713, L79.
\bibitem[\protect\citeauthoryear{Kreidberg et al.}{2015}]{kreidberg2015} Kreidberg, L., Line, M., Bean, J., et al., 2015, ApJ, 814, 66.
\bibitem[\protect\citeauthoryear{Kreidberg}{2015}]{batman2015} Kreidberg, L., 2015, arXiv:1507.08285.
\bibitem[\protect\citeauthoryear{Mandel \& Agol}{2002}]{ma2002} Mandel, K \& Agol, E., 2002, ApJ,
580, L171.
\bibitem[\protect\citeauthoryear{Ricker et al.}{2014}]{ricker2014} Ricker et al., 2014, SPIE, 9143.
\bibitem[\protect\citeauthoryear{Rowe et al.}{2015}]{rowe2015} Rowe et al., 2015, ApJS, 217, 16.
